Minimizing post-shock forecasting error using disparate information

### 2020-08-05 manuscript/manuscript.tex
changes by deletion of covariates and revision by comments

% use packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{url}
\usepackage{authblk}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{caption}
\usepackage{float}
\usepackage{tikz}
\usepackage{multirow}
\usepackage[linesnumbered, ruled,vlined]{algorithm2e}
% margin setup
\geometry{margin=0.8in}

@@

In this article, we will assume a simple auto regressive data generating process similar to that in \citet{blundell1998initial} with a general random effects structure. Therefore, our methodology is similar to the ``$K$ latent pooling'' framework of \cite{ramaswamy1993empirical}. However, our model formulation is more general than \cite{ramaswamy1993empirical} and our donor pool is formed from independent time series.
%The main idea of our methodology is to provide a scalar adjustment, based on estimated shock effects from the disparate time series, which is added to the original forecast at the known shock time point. 
We consider three aggregation techniques: simple averaging, inverse-variance weighted averaging, and similarity weighting. The latter technique is similar to the weighting in synthetic control methodology \citep{abadie2010synthetic}. Our auto regressive model will consider present day and lagged covariates to better motivate similarity weighting. The considered adjustment strategies all target the mean of the shock effect distribution. However, such an estimation strategy can reduce mean squared error (MSE) when variation in the shock effect distribution is small relative to the mean. We provide risk-reduction propositions that detail the conditions when the adjusted forecasts will work better than the original forecast. The involved parameters in the risk-reduction propositions and risk-reduction quantities can be estimated by a residual bootstrap procedure that we develop within. %These propositions and risk-reduction quantities are powerful tools for users in the sense that we can empirically aid a prospective evaluation about whether the three aggregation techniques will work well before the post-shock response is observed. 
We also motivate a simple leave-one-out cross validation procedure which can prospectively assess the performance of our shock effect adjustment estimators. This prospective assessment does not require the observation of the post-shock response. Our Monte Carlo simulation results show that the risk-reduction propositions are nearly perfectly correct when the model for the shock effects is identified well with appropriate covariates under a fixed design. We demonstrate the utility of our methodology in a real data analysis in which we forecast the stock price of Conoco Phillips shares that experienced a large structural shock on March 9th, 2020. We show that our proposed adjustment estimators yield much better results than no adjustment in this setting. We also use this example to demonstrate settings in which the shock effect may be decomposed into separate estimable parts. We now motivate our framework for post-shock forecasting.

@@

@@

\begin{figure}
  \begin{center}
    \includegraphics[height = 8cm]{comp.pdf}
    \caption{The time series experience a shock at $T_1^*+1=126$ with true shock effect $\alpha = 4.88$. The figure is a comparison between forecast without considering shock effects and the one uses simple averaging given $n=40$ disparate time series, and that the shock time is at $T_1^* +1=126$. The magenta dots represent least square estimate $\hat{\alpha}_i$ from disparate time series. The prediction of $\hat{y}^{2}_{T_1^*+1}$ and $\hat{y}^{1}_{T_1^*+1}$ differs only by an adjustment $\hat{\alpha}=5.22$. It is clear that $\hat{y}^{2}_{T_1^*+1}$ performs better than $\hat{y}^{1}_{T_1^*+1}$.}\label{figure1}
  \end{center}  
  \vspace{-.6cm}
\end{figure}

@@

The model $\mc{M}_1$ is defined as
\begin{align}
\mc{M}_1 \colon y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t}+ \beta_i'\mbf{x}_{i, t-1} + \varepsilon_{i,t}\label{equation1}
\end{align}
for $t = 1,\ldots,T_i$ and $i = 1,\ldots, n+1$, where $D_{i,t} = I(t = T_i^* + 1)$, 
$T_i^* < T_i$ and $\x_{i,t} \in \R^{p}$, $p \geq 1$.  We assume that the 
$\mbf{x}_{i,t}$'s are fixed and $T_i^*$s are known. The random effects structure for $\mc{M}_1$ is:
\begin{align*}
  \eta_i &\simiid  \eta ,\where \E{\eta} = 0, \var{\eta} = \sigma^2_{\eta}, \qquad i = 1, \ldots, n+1,\\
  \phi_i &\simiid \phi, \where |\phi|<1, \qquad i = 1, \ldots, n+1, \\
   \theta_i &\simiid \theta, \where \E{\theta}=\mu_{\theta}, \var{\theta}=\Sigma_{\theta}^2, \qquad i = 1, \ldots, n + 1, \\
\beta_i &\simiid \beta, \where \E{\beta}=\mu_{\beta}, \var{\beta}=\Sigma_{\beta}^2, \qquad i = 1, \ldots, n+1,\\
\alpha_i &\simiid \alpha, \where \E{\alpha}=\mu_{\alpha}, \var{\alpha}=\sigma_{\alpha}^2, \qquad  i = 1, \ldots, n+1; \\
\varepsilon_{i,t} &\simiid \normal{0}{\sigma^2}, \qquad t=1, \ldots, T_i, \; i = 1, \ldots, n+1,\\
\eta &\indep  \alpha_i \indep \phi \indep \theta \indep \varepsilon_{i,t}.
\end{align*}
Notice that $\mc{M}_1$ assumes that $\alpha_i$ are iid with $\E{\alpha_i}=\mu_{\alpha}$ 
for $i = 1, \ldots, n+1$.  %Although synthesizing information from disparate time series assumes similarities of shock effects, shock effects may differ in terms of their means in practice. 
We also consider a model where the shock effects are linear functions of covariates and 
lagged covariates with an additional additive mean-zero error.
%Additionally, $\mc{M}_1$ can be further improved in the sense that the shock effects may depend on the covariates around the shock time points, i.e., $T_i^*$ for $i = 1, \ldots, n+1$. 
%$\mc{M}_2$ can be constructed to accommodate those issues by imposing additional structures on the means of $\alpha_i$ as in (\ref{model2}) for $i = 1, \ldots, n+1$.
The random effects structure for this model (model $\mc{M}_2$) is:
\begin{align}
\mc{M}_2 \colon \begin{array}{l}
  y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t}+ \beta_i'\mbf{x}_{i, t-1} + \varepsilon_{i,t}\\[.2cm]
  \; \alpha_i = \mu_{\alpha}+\delta_{i}'\mbf{x}_{i, T_i^*+1}+\gamma_i'\mbf{x}_{i, T^*_i}+\t{\varepsilon}_{i},
\end{array}\label{model2}
\end{align}
for $i = 1, \ldots, n+1$, where the added random effects are
\begin{align*}
\t{\varepsilon}_{i} &\simiid  \E{\t{\varepsilon}}=0, \var{\t{\varepsilon}}=\sigma^2_{\alpha}, \qquad i = 1, \ldots, n+1,\\
\eta &\indep  \alpha_i \indep \phi \indep \theta \indep \varepsilon_{i,t} \indep \t{\varepsilon}_{i}.
\end{align*}
%and $\delta_i$ and $\gamma_i$ will be discussed shortly. 
We further define 
$\tilde{\alpha}_i=\mu_{\alpha}+\delta_i'\mbf{x}_{i, T_i^*+1}+\gamma_i'\mbf{x}_{i, T_i^*}$. 
We will investigate the post-shock aggregated estimators in $\mc{M}_2$ 
in settings where $\delta_i$ and $\gamma_i$ are either fixed or random. 
We let $\mc{M}_{21}$ denote model $\mc{M}_{2}$ with $\gamma_i = \gamma$ 
and $\delta_i = \delta$ for $i= 1, \ldots, n+1$, 
where $\gamma$ and $\delta$ are fixed unknown parameters.
We let $\mc{M}_{22}$ denote model $\mc{M}_{2}$ with the following random effects 
structure for $\gamma_i$ and $\delta_i$:
\begin{align*}
\begin{array}{c}
  \gamma_i \overset{iid}{\sim} \mrm{E}(\gamma) = \mu_\gamma, \var{\gamma} = \Sigma_\gamma \\
  \delta_i \overset{iid}{\sim} \mrm{E}(\delta) = \mu_\delta, \var{\delta} = \Sigma_\delta
\end{array}
   \quad \text{ with } \quad  \delta_i  \indep \t{\varepsilon}_{i} \quad  \text{ and } \quad \gamma_i  \indep \t{\varepsilon}_{i}.
\end{align*}
Note that $\delta_i$ and $\gamma_i$ may be dependent. We further define the parameter sets
\begin{align}
  \begin{array}{lll}
     \Theta &= &\{(\eta_i, \phi_i, \theta_i, \beta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i, \gamma_i)\colon t= 1, \ldots, T_i, i = 2, \ldots, n +1\},\\
    \Theta_1 &= &\{(\eta_i, \phi_i, \theta_i, \beta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i, \gamma_i)\colon t= 1, \ldots, T_i, i = 1\},\label{parameter}
  \end{array}
\end{align}
where $\Theta$ and $\Theta_1$ can adapt to $\mc{M}_1$ by dropping $\delta_i$ and 
$\gamma_i$. We assume this for notational simplicity.

\subsection{Forecast}
\label{forecast}
In this section we show how post-shock aggregate estimators improve upon standard 
forecasts that do not account for the shock effect.
%The interest of this study lies in comparing how consideration of shock effects 
%improves the prediction. 
More formally, we will consider the following candidate forecasts: 
\begin{align*}
  &\text{Forecast 1}: \hat y_{1,T_1^*+1}^1 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    + \hat\beta_1'\x_{1,T_1^*}, \\
  &\text{Forecast 2}: \hat y_{1,T_1^*+1}^2 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    + \hat\beta_1'\x_{1,T_1^*} + \hat{\alpha},
\end{align*}
where $\hat\eta_1$, $\hat\phi_1$, $\hat\theta_1$, and $\hat\beta_1$ are all ordinary least squares (OLS) estimators of $\eta_1$, $\phi_1$, $\theta_1$, and $\beta_1$ respectively, and $\hat{\alpha}$ is some form of estimator for the shock effect of time series of interest, i.e., $\alpha_1$. 
The first forecast ignores the presence of $\alpha_1$ while the second forecast 
incorporates an estimate of $\alpha_1$ that is obtained from the other independent forecasts under study. 
%Under $\mc{M}_1$ (Section \ref{modelsetup}), $\E{\alpha_1} = \mu_{\alpha}$.

Note that the two forecasts do not differ in their predictions for 
$y_{1,t}$, $t = 1,\ldots T_1^*$. Instead, they only differ in predicting 
$y_{1,T_1^*+1}$. Throughout the rest of this article we show that the collection of 
disparate time series $\{y_{i,t}, t = 2,\ldots,T_i, i = 1,\ldots,n\}$ has 
the potential to improve the forecasts for $y_{1, t}$ when $t > T_1^*$ under different 
circumstances for the dynamic panel model $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$. We specifically focus on predictions for $y_{1,T_1^*+1}$, the first post-shock response. It is important to note that in general $\hat{\alpha}$ 
is not a consistent estimator of the unobserved $\alpha_1$ nor does it converge 
to $\alpha_1$.  Despite these inferential shortcomings, adjustment of the forecast 
for $y_{1,T_1^*+1}$ through the addition of $\hat{\alpha}$ has 
the potential to lower forecast risk under several conditions corresponding to 
different estimators of $\alpha_1$. % which will be discussed shortly.

@@

We now construct the aggregate estimators of the shock effects that appear in 
Forecast 2 (see Section \ref{forecast}). We use these to forecast response values $y_{1, T_1^*+1}$ assuming that $T_1^*$ is known.  %Notice that the shock effect 
%for time series $1$ is typically unobserved. But we can gather knowledge from 
%disparate time series to construct some sort of estimators that may be helpful for 
%estimating $\E{\alpha_1}$. 
First, we introduce the procedures of parameter estimation for 
$\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ (see Section \ref{modelsetup}). 
Conditional on all regression parameters, previous responses, and covariates, 
the response variable $y_{i,t}$ in $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ 
has distribution 
$$
  y_{i,t} 
    \sim N(\eta_i + \alpha_iD_{i,t} + \phi_i y_{i,t-1} + \theta_i'\x_{i,t} 
      + \beta_i'\x_{i,t-1}, \sigma^2).
$$
For $i = 2, \ldots, n$, all parameters in this model will be estimated with ordinary least squares 
(OLS) using historical data of $t = 1, \ldots, n_i$. For $i = 1$, we estimate all the parameters but $\alpha_1$ using OLS procedures for $t=1, \ldots, T_1^*$. In particular, let $\hat{\alpha}_i$, $i = 2, \ldots, n+1$ be the OLS estimate 
of $\alpha_i$.  Note that parameter estimation 
for $\mc{M}_1$ is identically the same as that for $\mc{M}_{21}$ or $\mc{M}_{22}$.  We emphasize that $\alpha_i$s are random variables but the OLS estimation is conditioned on the their realizations from some distribution.

@@

We also consider the \emph{inverse-variance weighted estimator} 
in practical settings where the $T_i$'s and $T_i^*$'s vary greatly across $i$. 
The inverse-variance weighted estimator is defined as 
\begin{align*}
  \hat{\alpha}_{\rm IVW} = \frac{\sum_{i=2}^{n+1} \hat{\alpha}_i / \hat{\sigma}_{i\alpha}^2}{\sum_{i=2}^{n+1} 1/\hat{\sigma}_{i\alpha}^2},
  \quad \text{ where } \quad  \hat{\sigma}_{i\alpha}^2 = \hat{\sigma}^2_i( \mathbf{U}_i'\mbf{U}_i)_{22}^{-1},
\end{align*}
where  $\hat{\alpha}_i$ is the OLS estimator of $\alpha_i$, 
$\hat{\sigma}_i$ is the residual standard error from OLS estimation, 
and $\mbf{U}_i$ is the design matrix for OLS with respect to time series 
for $i = 2, \ldots, n+1$. Note that since $\sigma$ is unknown, estimation 
is required and the numerator and denominator terms are dependent in general. 
%It is hard to evaluate its expectation and variance. 
%But it is clear 
%that $\hat{\alpha}_{\rm IVW}$ will generally not be unbiased. 
However, $\hat{\alpha}_{\rm IVW}$ can be a reasonable estimator in 
practical settings. %when the change 
%point is not symmetric and/or there are some series with small or large number 
%of time points recorded. We will then investigate the performance of this 
%estimator in simulation (Section \ref{simulation}).
We do not provide closed form expressions for $\E{\hat{\alpha}_{\rm IVW}}$ 
and $\var{\hat{\alpha}_{\rm IVW}}$ but empirical performance of 
$\hat{\alpha}_{\rm IVW}$ is assessed via Monte Carlo simulation 
(see Section \ref{simulation}).

@@

 \begin{enumerate}[label = (\roman*)]
    \item Under $\mc{M}_{1}$, $\E{\hat{\alpha}_{\rm adj}}=\E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha}$.
    \item Under $\mc{M}_{21}$, 
    \begin{align*}
      \E{\hat{\alpha}_{\rm adj}} = \mu_{\alpha} + \frac{1}{2} \sum_{i=2}^{n+1} \delta' \mbf{x}_{i, T_i^*+1} + \frac{1}{n} \sum_{i=2}^{n+2} \gamma' \mbf{x}_{i, T_i^*} 
      \quad \text{ and } \quad 
       \E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha} + \delta'\mbf{x}_{1, T_1^*+1} + \gamma' \mbf{x}_{1, T_1^*}.
    \end{align*}
    \item Under $\mc{M}_{22}$,
    \begin{align*}
      \E{\hat{\alpha}_{\rm adj}} = \mu_{\alpha} + \frac{1}{2} \sum_{i=2}^{n+1} \mu_{\delta}' \mbf{x}_{i, T_i^*+1} + \frac{1}{n} \sum_{i=2}^{n+2}  \mu_{\gamma}'\mbf{x}_{i, T_i^*} 
      \quad \text{ and } \quad 
       \E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha} + \mu_{\delta}'\mbf{x}_{1, T_1^*+1} + \mu_{\gamma}' \mbf{x}_{1, T_1^*}.
    \end{align*}
  \end{enumerate}
Formal justification for these results can be found in Appendix. Note that $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ are not unbiased estimators for $\alpha_1$. Notice that under $\mc{M}_{1}$, $\hat{\alpha}_{\rm adj}$ and $\hat{\alpha}_{\rm adj}$ are unbiased estimators for $\E{\alpha_1}=\mu_{\alpha}$ (see distributional details of $\alpha_1$ in Section \ref{modelsetup}). Nevertheless, $\hat{\alpha}_{\rm adj}$ is a biased estimator for $\E{\alpha_1}$ but $\hat{\alpha}_{\rm wadj}$ is an unbiased estimator for $\E{\alpha_1}$ under both $\mc{M}_{21}$ and $\mc{M}_{22}$. Thus, we collect these results  as the following proposition.

>>>>

% use packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{url}
\usepackage{authblk}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{caption}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{tikz}
\usepackage{multirow}
\usepackage[linesnumbered, ruled,vlined]{algorithm2e}
% margin setup
\geometry{margin=0.8in}

@@

In this article, we will assume a simple auto regressive data generating process similar to that in \citet{blundell1998initial} with a general random effects structure. Therefore, our methodology is similar to the ``$K$ latent pooling'' framework of \cite{ramaswamy1993empirical}. However, our model formulation is more general than \cite{ramaswamy1993empirical}. In our model, the donor pool can consist of dependent time series but time series within the donor pool should be independent of the time series of interest. However, mutual independence among time series in the donor pool can aid prospective evaluation of the reliability of our method. 
%The main idea of our methodology is to provide a scalar adjustment, based on estimated shock effects from the disparate time series, which is added to the original forecast at the known shock time point. 
We consider three aggregation techniques: simple averaging, inverse-variance weighted averaging, and similarity weighting. The latter technique is similar to the weighting in synthetic control methodology \citep{abadie2010synthetic}. Our auto regressive model will consider present day and lagged covariates to better motivate similarity weighting. The considered adjustment strategies all target the mean of the shock effect distribution. However, such an estimation strategy can reduce mean squared error (MSE) when variation in the shock effect distribution is small relative to the mean. We provide risk-reduction propositions that detail the conditions when the adjusted forecasts will work better than the original forecast. The involved parameters in the risk-reduction propositions and risk-reduction quantities can be estimated by a residual bootstrap procedure that we develop within. %These propositions and risk-reduction quantities are powerful tools for users in the sense that we can empirically aid a prospective evaluation about whether the three aggregation techniques will work well before the post-shock response is observed. 
We also motivate a simple leave-one-out cross validation procedure which can prospectively assess the performance of our shock effect adjustment estimators. This prospective assessment does not require the observation of the post-shock response. Our Monte Carlo simulation results show that the risk-reduction propositions are nearly perfectly correct when the model for the shock effects is identified well with appropriate covariates under a fixed design. We demonstrate the utility of our methodology in a real data analysis in which we forecast the stock price of Conoco Phillips shares that experienced a large structural shock on March 9th, 2020. We show that our proposed adjustment estimators yield much better results than no adjustment in this setting. We also use this example to demonstrate settings in which the shock effect may be decomposed into separate estimable parts. We now motivate our framework for post-shock forecasting.

@@

\begin{figure}[H]
  \begin{center}
    \subfloat[A comparison between forecast without considering shock effects and the one uses simple averaging given $n=40$ disparate time series, and that the shock time is at $T_1^* +1=126$. The magenta dots represent least square estimate $\hat{\alpha}_i$ from disparate time series. The prediction of $\hat{y}^{2}_{T_1^*+1}$ and $\hat{y}^{1}_{T_1^*+1}$ differs only by an adjustment $\hat{\alpha}=5.22$. It is clear that $\hat{y}^{2}_{T_1^*+1}$ performs better than $\hat{y}^{1}_{T_1^*+1}$.]{\includegraphics[height = 7cm]{comp.pdf}}\\
     \subfloat[Histogram for $\hat{\alpha}_i$ for $i = 2, \ldots, n+1$]{\includegraphics[height = 7cm]{alphahist.pdf}}
  \end{center}
     \caption{The time series experience a shock at $T_1^*+1=126$ with true shock effect $\alpha = 4.21$. (a) presents the comparison of the prediction without adjustment and one with. (b) shows the histogram for the $\alpha_i$s used in estimating shock $\alpha_1$.}\label{figure1}
\end{figure}

@@

@@

Let $I(\cdot)$ be an indicator function, $T_i$ be the time length of the time series $i$ for $i = 1, \ldots, n+1$, and $T_i^*$ be the time point just before the one when the shock is \emph{known} to occur, with $T_i^* < T_i$.  For $t= 1, \ldots, T_i$ and $i = 1, \ldots, n+1$, the model $\mc{M}_1$ is defined as
\begin{align}
\mc{M}_1 \colon y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\label{equation1}
\end{align}
 where $D_{i,t} = I(t = T_i^* + 1)$ 
and $\x_{i,t} \in \R^{p}$ with $p \geq 1$.  We assume that the 
$\mbf{x}_{i,t}$'s are fixed. Let $|x|$ denote the absolute value of $x$ for $x\in \reals$. For $i = 1, \ldots, n+1$ and $t=1, \ldots, T_i$, the random effects structure for $\mc{M}_1$ is:
\begin{align*}
  \eta_i &\simiid  \; \E{\eta_i} = 0, \var{\eta_i} = \sigma^2_{\eta}\\
  \phi_i &\simiid \; |\phi_i|<1, \\
   \theta_i &\simiid  \; \E{\theta_i}=\mu_{\theta}, \var{\theta_i}=\Sigma_{\theta}^2 \\
\alpha_i &\simiid \; \E{\alpha_i}=\mu_{\alpha}, \var{\alpha_i}=\sigma_{\alpha}^2, \\
\varepsilon_{i,t} &\simiid \; \E{\varepsilon_{i,t}} =0, \var{\varepsilon_{i,t}}=\sigma^2 \where \sigma > 0 ,  \\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t}.
\end{align*}
Notice that $\mc{M}_1$ assumes that $\alpha_i$ are iid with $\E{\alpha_i}=\mu_{\alpha}$ 
for $i = 1, \ldots, n+1$.  %Although synthesizing information from disparate time series assumes similarities of shock effects, shock effects may differ in terms of their means in practice. 
We also consider a model where the shock effects are linear functions of covariates with an additional additive mean-zero error.
%Additionally, $\mc{M}_1$ can be further improved in the sense that the shock effects may depend on the covariates around the shock time points, i.e., $T_i^*$ for $i = 1, \ldots, n+1$. 
%$\mc{M}_2$ can be constructed to accommodate those issues by imposing additional structures on the means of $\alpha_i$ as in (\ref{model2}) for $i = 1, \ldots, n+1$.
For $i = 1, \ldots, n+1$, the random effects structure for this model (model $\mc{M}_2$) is:
\begin{align}
\mc{M}_2 \colon \begin{array}{l}
  y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\\[.2cm]
  \; \alpha_i = \mu_{\alpha}+\delta_{i}'\mbf{x}_{i, T_i^*+1}+ \t{\varepsilon}_{i},
\end{array}\label{model2}
\end{align}
 where the added random effects are
\begin{align*}
\t{\varepsilon}_{i} &\simiid  \E{\t{\varepsilon}}=0, \var{\t{\varepsilon}}=\sigma^2_{\alpha} \where \sigma_{\alpha}>0\\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t} \indep \t{\varepsilon}_{i}.
\end{align*}
%and $\delta_i$ and $\gamma_i$ will be discussed shortly. 
We further define 
$\tilde{\alpha}_i=\mu_{\alpha}+\delta_i'\mbf{x}_{i, T_i^*+1}$. 
We will investigate the post-shock aggregated estimators in $\mc{M}_2$ 
in settings where $\delta_i$ is either fixed or random. 
We let $\mc{M}_{21}$ denote model $\mc{M}_{2}$ with $\delta_i = \delta$ for $i= 1, \ldots, n+1$, 
where $\delta$ is a  fixed unknown parameter.
We let $\mc{M}_{22}$ denote model $\mc{M}_{2}$ with the following random effects 
structure for $\delta_i$:
\begin{align*}
\delta_i \overset{iid}{\sim}\;  \mrm{E}(\delta) = \mu_\delta, \var{\delta} = \Sigma_\delta
   \quad \text{ with } \quad  \delta_i  \indep \t{\varepsilon}_{i}.
\end{align*}
We further define the parameter sets
\begin{align}
  \begin{array}{lll}
     \Theta &= &\{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon t= 1, \ldots, T_i, i = 2, \ldots, n +1\},\\
    \Theta_1 &= &\{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon t= 1, \ldots, T_i, i = 1\},\label{parameter}
  \end{array}
\end{align}
where $\Theta$ and $\Theta_1$ can adapt to $\mc{M}_1$ by dropping $\delta_i$. We assume this for notational simplicity.

\subsection{Forecast}
\label{forecast}
In this section we show how post-shock aggregate estimators improve upon standard 
forecasts that do not account for the shock effect.
%The interest of this study lies in comparing how consideration of shock effects 
%improves the prediction. 
More formally, we will consider the following candidate forecasts: 
\begin{align*}
  &\text{Forecast 1}: \hat y_{1,T_1^*+1}^1 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    , \\
  &\text{Forecast 2}: \hat y_{1,T_1^*+1}^2 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    + \hat{\alpha},
\end{align*}
where $\hat\eta_1$, $\hat\phi_1$, and $\hat\theta_1$ are all ordinary least squares (OLS) estimators of $\eta_1$, $\phi_1$, and $\theta_1$, respectively, and $\hat{\alpha}$ is some form of estimator for the shock effect of time series of interest, i.e., $\alpha_1$. 
The first forecast ignores the presence of $\alpha_1$ while the second forecast 
incorporates an estimate of $\alpha_1$ that is obtained from the other independent forecasts under study. 
%Under $\mc{M}_1$ (Section \ref{modelsetup}), $\E{\alpha_1} = \mu_{\alpha}$.

Note that the two forecasts do not differ in their predictions for 
$y_{1,t}$, $t = 1,\ldots T_1^*$. Instead, they only differ in predicting 
$y_{1,T_1^*+1}$. Throughout the rest of this article we show that the collection of 
disparate time series $\{y_{i,t}, t = 1,\ldots,T_i, i = 2,\ldots,n+1\}$ has 
the potential to improve the forecasts for $y_{1, t}$ when $t > T_1^*$ under different 
circumstances for the dynamic panel model $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$. Improvement will be  measured by \emph{risk-reduction quantity}, which is the difference between the risk of Forecast 1 and Forecast 2. We will return to the theoretical details of risk-reduction quantity in Section \ref{properties}.

We specifically focus on predictions for $y_{1,T_1^*+1}$, the first post-shock response. It is important to note that in general $\hat{\alpha}$ 
is not a consistent estimator of the unobserved $\alpha_1$ nor does it converge 
to $\alpha_1$.  Despite these inferential shortcomings, adjustment of the forecast 
for $y_{1,T_1^*+1}$ through the addition of $\hat{\alpha}$ has 
the potential to lower forecast risk under several conditions corresponding to 
different estimators of $\alpha_1$. % which will be discussed shortly.

@@

We now construct the aggregate estimators of the shock effects that appear in 
Forecast 2 (see Section \ref{forecast}). We use these to forecast response values $y_{1, T_1^*+1}$ assuming that $T_1^*$ is known.  %Notice that the shock effect 
%for time series $1$ is typically unobserved. But we can gather knowledge from 
%disparate time series to construct some sort of estimators that may be helpful for 
%estimating $\E{\alpha_1}$. 
First, we introduce the procedures of parameter estimation for 
$\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ (see Section \ref{modelsetup}). 
Conditional on all regression parameters, previous responses, and covariates, 
the response variable $y_{i,t}$ in $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ 
has distribution 
$$
  y_{i,t} 
    \sim N(\eta_i + \alpha_iD_{i,t} + \phi_i y_{i,t-1} + \theta_i'\x_{i,t} 
      , \sigma^2).
$$
For $i = 2, \ldots, n+1$, all parameters in this model will be estimated with ordinary least squares 
(OLS) using historical data of $t = 1, \ldots, T_i$. For $i = 1$, we estimate all the parameters but $\alpha_1$ using OLS procedures for $t=1, \ldots, T_1^*$. In particular, let $\hat{\alpha}_i$, $i = 2, \ldots, n+1$ be the OLS estimate 
of $\alpha_i$.  Note that parameter estimation 
for $\mc{M}_1$ is identically the same as that for $\mc{M}_{21}$ or $\mc{M}_{22}$.  We emphasize that $\alpha_i$s are random variables but the OLS estimation is conditioned on the their realizations from some distribution.

@@

We also consider the \emph{inverse-variance weighted estimator} 
in practical settings where the $T_i$'s and $T_i^*$'s vary greatly across $i=2, \ldots, n+1$. 
The inverse-variance weighted estimator is defined as 
\begin{align*}
  \hat{\alpha}_{\rm IVW} = \frac{\sum_{i=2}^{n+1} \hat{\alpha}_i / \hat{\sigma}_{i\alpha}^2}{\sum_{i=2}^{n+1} 1/\hat{\sigma}_{i\alpha}^2},
  \quad \text{ where } \quad  \hat{\sigma}_{i\alpha}^2 = \hat{\sigma}^2_i( \mathbf{U}_i'\mbf{U}_i)_{22}^{-1},
\end{align*}
where  $\hat{\alpha}_i$ is the OLS estimator of $\alpha_i$, 
$\hat{\sigma}_i$ is the residual standard error from OLS estimation, 
and $\mbf{U}_i$ is the design matrix for OLS with respect to time series 
for $i = 2, \ldots, n+1$. Note that since $\sigma$ is unknown, estimation 
is required and the numerator and denominator terms are dependent in general. 
%It is hard to evaluate its expectation and variance. 
%But it is clear 
%that $\hat{\alpha}_{\rm IVW}$ will generally not be unbiased. 
However, $\hat{\alpha}_{\rm IVW}$ can be a reasonable estimator in 
practical settings. %when the change 
%point is not symmetric and/or there are some series with small or large number 
%of time points recorded. We will then investigate the performance of this 
%estimator in simulation (Section \ref{simulation}).
We do not provide closed form expressions for $\E{\hat{\alpha}_{\rm IVW}}$ 
and $\var{\hat{\alpha}_{\rm IVW}}$ but empirical performance of 
$\hat{\alpha}_{\rm IVW}$ is assessed via Monte Carlo simulation 
(see Section \ref{simulation}).

@@

 \begin{enumerate}[label = (\roman*)]
    \item Under $\mc{M}_{1}$, $\E{\hat{\alpha}_{\rm adj}}=\E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha}$.
    \item Under $\mc{M}_{21}$, 
    \begin{align*}
      \E{\hat{\alpha}_{\rm adj}} = \mu_{\alpha} + \frac{1}{n} \sum_{i=2}^{n+1} \delta' \mbf{x}_{i, T_i^*+1} 
      \quad \text{ and } \quad 
       \E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha} + \delta'\mbf{x}_{1, T_1^*+1} .
    \end{align*}
    \item Under $\mc{M}_{22}$,
    \begin{align*}
      \E{\hat{\alpha}_{\rm adj}} = \mu_{\alpha} + \frac{1}{n} \sum_{i=2}^{n+1} \mu_{\delta}' \mbf{x}_{i, T_i^*+1}
      \quad \text{ and } \quad 
       \E{\hat{\alpha}_{\rm wadj}} = \mu_{\alpha} + \mu_{\delta}'\mbf{x}_{1, T_1^*+1} .
    \end{align*}
  \end{enumerate}
Formal justification for these results can be found in Appendix. Note that $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ are not unbiased estimators for $\alpha_1$. Notice that under $\mc{M}_{1}$, $\hat{\alpha}_{\rm adj}$ and $\hat{\alpha}_{\rm adj}$ are unbiased estimators for $\E{\alpha_1}=\mu_{\alpha}$ (see distributional details of $\alpha_1$ in Section \ref{modelsetup}). Nevertheless, $\hat{\alpha}_{\rm adj}$ is a biased estimator for $\E{\alpha_1}$ but $\hat{\alpha}_{\rm wadj}$ is an unbiased estimator for $\E{\alpha_1}$ under both $\mc{M}_{21}$ and $\mc{M}_{22}$. Thus, we collect these results  as the following proposition.
